### 2007-09-12 doc/a-rex/arex_tech_doc.tex
Changed title on request of Balazs. Language fixes from Adrian.

\begin{center}
\textsc{\large The ARC Computational Job Management Module - A-Rex}
\par\end{center}{\large \par}

@@

Each job gets a directory on the CE called the \emph{session directory}
(SD). Input files are gathered in the SD. The job is supposed to produce
new data files also in the SD. GM does not guarantee the availability
of any other places accessible by the job other than SD (unless such
place is part of requested Runtime Environement). The SD is also the
only place which is controlled by the GM. It is is accessible by the
user from outside through HTTPS and GridFTP protocols. Any file created
outside the SD is not controlled by the GM. Any exchange of data between
client and GM (including also program modules) is done through GridFTP
\cite{gridftp} or HTTPS protocols depending on selected interface.
A URL for accessing input/output files is obtained through the Information
System or WS Information Interface of A-Rex.

@@

Each job is initiated and controlled through GFS or WS. Complete description
of job (not data) is passed to the GM through GFS in RSL \cite{rsl}
or through WS in JSDL-coded \cite{jsdl} description (job description
- JD).

@@

The main task of the GM is to take care of processing input and output
data (files) of the job. Input files are gathered in SD. There are
2 ways to put file into the SD:

\begin{itemize}
\item Downloads initiated by the GM. Such files (name and source) are defined
in the JD. It is a sole responsibility of the GM to make sure that
a file will be available in the SD.\\
The supported sources are at the moment: GridFTP, FTP, HTTP, HTTPS
(HTTP over SSLv3) and HTTPg (HTTP over GSI). Also some nonstandard
sources are supported. Those are described below.
\item Upload initiated by the user directly or through the User Interface
(UI). Because the SD becomes available immediately at the time of
submission of JD, UI can (and should) use that to upload data files
which are not otherwise accessible by the GM. An example of such files
can be the main executable of the job, files containing job's options/parameters,
etc. These files can (and should) also be specified in the JD.
\end{itemize}
\textbf{There is no} other reliable way for a job to obtain input
data on the CE belonging to NorduGrid. Access to AFS, NFS, FTP, HTTP
and any other remote data transport during execution of the job is
not guaranteed (at least not by GM).

Job stores output files in the SD. Those files also belong to 2 groups:

\begin{itemize}
\item Files which are supposed to be moved to a \emph{Storage Element} (SE)
and optionally registered in some \emph{Indexing Service} like \emph{Replica
Catalog} (RC). The GM takes care of those files. They have to be specified
in the JD. If job fails during any stage of processing no attept is
taken to transfer those files to their final destination, unless there
is option \emph{preserve=yes} specified in their URLs.
\item Files which are supposed to be fetched by the user. The user runs
UI to obtain those files. They \textbf{must} also be specified in
the JD.
\end{itemize}

@@

\caption{\label{job states diagram}Job states}
\end{figure}
A user can examine the state of a job by querying the NorduGrid Information
System (IS) using the UI or any other tool. Please remember that IS
can manipulate with state names to make them more user friendly and
to combine them with states introduced by other parts of whole setup.
Another way is to access virtual informational files through GridFTP
interface or to use query metod of JCS.

Configuration can put limits on amout of simultaneous jobs at some
states. If such limit is reached job stays in it's current state waiting
for free slot. This situation is presented by prepending current state
name with \textbf{PENDING:} status mark.

Below is description of all actions taken by the GM at every state:

\begin{itemize}
\item \textbf{Accepted} - At this state the job has been submitted to a
CE but not processed yet. The GM will analyze the JD and move to the
next stage. If JD can not be processed the job will be canceled and
moved to the state \textbf{Finishing}.
\item \textbf{Preparing} - The input data is being gathered in the SD. The
GM is downloading files specified in the JD and waiting for files
which are supposed to be downloaded by the UI. If all files are successfully
gathered the job moves to the next state. If \textbf{any} file can't
be downloaded or it takes UI too long to upload a file - the job moves
to \textbf{Finishing} state. It is possible to put limit on number
of simultaneous \textbf{Preparing} jobs. Those jobs out of limit will
stay in previous \textbf{Accepted} state with PENDING mark. Exceptions
are jobs which has no files to be downloaded. Those are processed
out of limits.
\item \textbf{Submitting} - This is a point of interaction with \emph{Local
Resource Management System} (LRMS). At the moment PBS is supported
best and correspoding backends are provided with default installation.
The job is being submitted for execution. If the local job submission
is successful the job moves to the next state. Otherwise it moves
to \textbf{Finishing}. It is possible to limit number of jobs in \textbf{Submitting}
and following \textbf{InLRMS} states.
\item \textbf{InLRMS} - The job is queued or being executed in the LRMS.
The GM takes no actions except waiting until job finishes.
\item \textbf{Cancelling} - Necessary action to cancel job in the LRMS is
being taken.
\item \textbf{Finishing} - The output data is being processed. Specified
data files are moved to the specified SEs and are optionally registered
at RC. The user can download data files from the SD by using UI or
any other tool. All the files not specified as output files are removed
from the SD at very beginning of this state. It is possible to limit
number of simultaneous jobs in this state.
\item \textbf{Finished} - No more processing is performed by the GM. The
user can continue to download data files from the SD. The SD is kept
available for some time (default is 1 week). After that it is moved
to the state \textbf{Deleted}. The 'deletion' time can be queried
at NorduGrid Information System as attribute \texttt{nordugrid-pbs-job-sessiondirerasetime}
of \texttt{nordugrid-pbs-job}. If job was moved to \textbf{Finished}
because of failure, it may be restarted on request of client. Job
is moved to a state previous to one which failed and is assigned mark
PENDING. This is needed in order to not break the configuration limits.
Exception is a job failed in \textbf{InLRMS} state and lacking input
files specified in JD. Such job is treated like failed in \textbf{Preparing}
state.
\item \textbf{Deleted} - Job is moved to this state if user does not request
job to be cleaned. Only minimal subset of information about such job
is kept.
\end{itemize}
In the case of the failure special processing is applied to output
files. All specified output files are treated as \textbf{downloadable
by user}. No files will be moved to the SE.

@@

The GM and it's components support following data transfer protocols
and corresponding URLs: \emph{ftp, gsiftp, http, httpg, https, se,
rc} and \emph{rls.} For more information please see {}``Protocols,
Uniform Resource Locators (URL) and extensions supported in ARC''
document \cite{urls}.

@@

For each local UNIX user listed in the GM configuration a \textit{control
directory} exists. In this directory the GM stores information about
jobs belonging to that user. Multiple users can share the same \textit{control
directory}. To make it easier to recover in the case of failure, the
GM stores most information in files rather than in memory. All files
belonging to same job have names starting with \textbf{job.ID.} here
ID is the job identifier.

@@

\begin{itemize}
\item \textit{job.ID.status} - current state of the job. It contains one
word of text representing the current state of the job. Possible values
are :

\begin{itemize}
\item ACCEPTED
\item PREPARING
\item SUBMITTING
\item INLRMS
\item FINISHING
\item FINISHED
\item CANCELING
\item DELETED
\end{itemize}
\end{itemize}
See section \ref{Section: Job Flow} for a description of the various
states. Additionally each value can be prepended with a prefix {}``PENDING:''
(like PENDING:ACCEPTED, see section \ref{Section: Job Flow}). That
is used to show that the job is \emph{ready} to be moved to a next
state and it stays in a current state \emph{only} because some limits
set in configuration are exceeded.

\begin{itemize}
\item \textit{job.ID.description} - contains the RSL description of the
job.
\item \textit{job.ID.local} - information about job used by the GM. It consists
of lines of format \textit{{}``name = value''} . Not all of them
are always available. The following names are defined:

\begin{itemize}
\item \textit{subject} - user subject also known as the distinguished name
(DN)
\item \textit{starttime} - the GMT time when the job was accepted represented
in Generalized Time format of LDAP 
\item \textit{lifetime} - time period to live for the SD after job finished
in seconds
\item \textit{cleanuptime} - the GMT time when job to be removed from cluster
and SD deleted in Generalized Time format
\item \textit{notify} - email addresses and flags to send mail to about
job specified status changes
\item \textit{processtime} - the GMT time when to start processing the job
in Generalized Time format
\item \textit{exectime} - the GMT time when to start job execution in Generalized
Time format
\item \textit{expiretime} - the GMT time when credentials delegated to job
expire in Generalized Time format
\item \textit{rerun} - number of retries left to run the job
\item \textit{jobname} - name of the job as supplied by the user
\item \textit{lrms} - name of LRMS to run the job at
\item \textit{queue} - name of the queue to run the job at
\item \textit{localid} - job id in LRMS (appears only then the job is at
state \textbf{InLRMS})
\item \textit{args} - list of command-line arguments including the executable
\item \textit{downloads} - number of files to download into SD before execution
\item \textit{uploads} - number of files to upload from SD after execution
\item \textit{gmlog} - directory name which holds files containing information
about job when accessed through GridFTP interface
\item \textit{clientname} - name and ip address:port of client machine (name
is provided by user interface)
\item \textit{clientsoftware} - version of software used to submit job
\item \textit{sessiondir} - SD of job
\item \textit{failedstate} - state at which job failed (available only if
it is possible to restart job)
\item \textit{jobreport} - URL of \emph{logger service} used to keep track
of executed jobs (one requested by user)
\end{itemize}
\end{itemize}
This file is filled partially during job submission and fully when
the job moves from the \textbf{Accepted} to the \textbf{Preparing}
state.

@@

The job submission and control interface is based on document produced
by OGF OGSA Basic Execution Services (BES) Working Group\cite{bes}.
The exchange of SOAP messages is performed over HTTPS protocol. The
BES interface is represented by two port-types - BES-Management and
BES-Factory. The former is made to control the A-REX service itself
and thus defines operations to start and stop the functionality of
the BES service. A-REX does not implement remote control of service
functionality. Hence the BES-Management port-type is not implemented.
The BES-Factory port-type provides operations to submit new jobs (to
create an activity in terms of BES) and to monitor its state. It also
has an ability to provide information about the service. A-REX fully
implements the functionality of this port-type.

@@

\begin{itemize}
\item \emph{NotAuthorizedFault}: Indicates that client is not allowed to
do this operation.
\item \emph{InvalidActivityIdentifierFault}: There is no such job/activity.
\item \emph{CantApplyOperationToCurrentStateFault}: Requested transition
is not possible.
\end{itemize}
\end{itemize}
\end{itemize}
On result of this command, the job should be put into the requested
state. If such a procedure cannot be performed immediately then the
corresponding sequence is initiated and fault OperationWillBeAppliedEventuallyFault
will be returned.

@@

\begin{itemize}
\item \emph{DelegatedTokenType DelegatedToken}: Contains an X509 proxy certificate
based on the public key from the DelegateCredentialsInit signed by
the user's proxy certificate. Also includes the Id element which identifies
the private key stored at the service side associated with these credentials.
The reference element refers to the object which these credentials
should be applied to in a way specific to the service. The same element
must also be used for delegating credentials as part of other operations
on service.
\end{itemize}
\item Output(s)

@@

\begin{lyxlist}{00.00.0000}
\item [{Filename}] - name of the data file on the executing node.
\item [{\emph{Source}}] - source where the file will be taken from before
execution.
\item [{\emph{Target}}] \emph{-} destination the file will be delivered
to after execution.
\end{lyxlist}
\item [{\emph{Input}\,\ (POSIX)}] - file to be used as standard input
for the executable.
\item [{\emph{Output}\ (POSIX)}] - file to be used as standard output
for the executable.
\item [{\emph{Error}\ (POSIX)}] - file to be used as standard error for
the executable.
\item [{\emph{MemoryLimit}\ (POSIX)}] - amount of physical memory needed
for execution.
\item [{\emph{TotalPhysicalMemory}}] - same as MemoryLimit.
\item [{\emph{IndividualPhysicalMemory}}] - same as MemoryLimit.
\item [{\emph{CPUTimeLimit}\ (POSIX)}] - maximal amount of CPU time needed
for execution.
\item [{\emph{TotalCPUTime}}] - same as CPUTimeLimit.
\item [{\emph{IndividualCPUTime}}] - same as CPUTimeLimit.
\item [{\emph{WallTimeLimit}\ (POSIX)}] - amount of clock time needed
for execution.
\item [{\emph{TotalCPUCount}}] \emph{-} number of CPUs needed for execution.
\item [{\emph{IndividualCPUCount}}] - same as TotalCPUCount.
\end{lyxlist}

@@

\begin{itemize}
\item \textit{list} - stores names of the files (8 digit numbers) and corresponding
URLs delimited by blank space. Each pair is delimited by some amount
of \textbackslash{}0 codes. Also creation and expiration times are
stored if available
\item \textit{old} - stores URLs which have been removed from cache. Records
are delimited by some amount of \textbackslash{}0 codes and are meant
to be removed by some external routine.
\item \textit{new} - stores URLs which have been added to cache. Records
are delimited by some amount of \textbackslash{}0 codes and are removed
when corresponding files are removed from cache. They can also be
handled by some external routines. Every time record is added to \emph{old}
it is removed from \emph{new}.
\item \textit{statistics} - consists of strings containing \emph{name=value}
. Following names are defined:

@@

If job requests input file, which can/allowed to be cached, it is
stored in cache directory instead and soft-link is created in the
SD, pointing to that file. Or file can be stored in cache and then
copied to the SD. Last option is more secure and hence advised.

Before downloading file the GM tries to determine it's size and to
preallocate space in cache directory, by writing file of same size.
If that fails (file system has no more space), it tries to remove
oldest cache files, which are not being used by any job. That means
\textbf{hard limit of cache size is space available at file-system}.
In case cache gets full and it is impossible to free any space, download
fails and is retried without using cache.

Before giving access to already cached file the GM contacts initial
file source to check if user is allowed to do that if protocols allows
to do that.

Also file creation or validity times are checked to make sure cached
file is fresh enough. If it is impossible to obtain creation and invalidation
times for file it is invalidated 24 hours after downloaded.

Also the GM checks cache periodically. If used space exceeds high
water-mark given in configuration file (\emph{softsize}) it tries
to remove oldest unused files to reduce size to low water-mark. This
sets soft limit of cache size.

There are 2 kinds of caches supported. Files in \textit{private} cache
are owned by Unix user to which grid user is mapped. Those files are
readable only by that particular Unix user. Another kind of cache
is \textit{shared}. Files are owned by Unix user who started GM and
are readable by everyone.

@@

The GM consists of few separate executable modules. Those are:

\begin{itemize}
\item \textit{grid-manager} - Main module. It is responsible for processing
the job, moving it through states, running other modules.
\item \textit{downloader} - This is a module responsible for gathering input
files in the SD. It processes the \textit{job.ID.input} file and updates
it.
\item \textit{uploader} - This module is responsible for delivering output
files to the specified SEs and registration at the RC. It processes
and updates the \textit{job.ID.output} file.
\item \emph{cache-register} - Utility to register cached data into Indexing
Services like RC and RLS. It reads and modifies cache informational
files \emph{old} and \emph{new} \ref{sec:cache}. Configuration is
read directly from the GM's configuration file \ref{SubSection: ConfigFile}.
It is run by the GM every 5 minutes.
\item \emph{frontend-info-collector} - Utility to gather information about
frontend and to put it into \emph{job.ID.diag} file.
\item \emph{gm-kick} - Sends signal to the GM though FIFO file to wake it
up. It's used to increase responsiveness of GM.
\end{itemize}
Following modules are always run under Unix account to which user
is mapped.

\begin{itemize}
\item \textit{smtp-send.sh} and \textit{smtp-send} - These are the modules
responsible for sending e-mail notifications to the user. The format
of the mail messages can be easily changed by editing the simple shell
script \textit{smtp-send.sh}. 
\item \textit{submit-{*}-job} - Here {*} stands for the name of the LRMS.
Curently supported LRMS are PBS/Torque, Condor and SGE. Also \emph{fork}
pseudo-LRMS is supported for testing purposes. This module is responsible
for the job submission to the LRMS.
\item \textit{cancel-{*}-job} - This one is for canceling the job, which
was submitted to LRMS.
\item \textit{scan-{*}-job} -This shell script is responsible for notifying
the GM about completion of the job. It's implementation for PBS system
uses server logs to extract information about jobs. If logs are not
available it uses less reliable \emph{qstat} command for that. other
backends use different techniques.
\end{itemize}
Also there is administrator utility:

\begin{itemize}
\item \textit{gm-jobs} - prints list of jobs available on cluster and amount
of jobs in every state.\\
\hspace*{0.5cm}gm-jobs {[}-h] {[}-l] {[}-u uid] {[}-U name]\\
\hspace*{0.5cm}-l-l - print more information about each job,\\
\hspace*{0.5cm}-l-u - pretend utility is run by user with id \emph{uid},\\
\hspace*{0.5cm}-l-l - pretend utility is run by user with name \emph{name}.
\end{itemize}
GM comes with plugins useable for various authorization purposes (see
for example description of \emph{authplugin} command below):

\begin{itemize}
\item \textit{inputcheck} - checks if all input files specified in job description
are downloadable.\\
\hspace*{0.5cm}inputcheck {[}-h] {[}-d debug\_level] RSL\_fle {[}proxy\_file]\\
\hspace*{0.5cm}-lRSL\_file -file with job description,\\
\hspace*{0.5cm}-lproxy\_file - credentials proxy.
\item \textit{lcas} - executes LCAS plugins on credentials and returns 0
if authorization passed.\\
\hspace*{0.5cm} lcas credentials description {[}library {[}db {[}directory]]]\\
\hspace*{0.5cm}-lcredentials - path to file with credentials to authorize,\\
\hspace*{0.5cm}-ldescription - path to file with job description,\\
\hspace*{0.5cm}-llibrary - path to LCAS library (full or relative
to LCAS directory),\\
\hspace*{0.5cm}-ldb - path to LCAS DB file (full or relative to LCAS
directory),\\
\hspace*{0.5cm}-ldirectory - LCAS directory.
\end{itemize}

>>>>

\begin{center}
\textsc{\large The ARC Computational Job Management Component
- A-Rex}
\par\end{center}{\large \par}

@@

Each job gets a directory on the CE called the \emph{session directory}
(SD). Input files are gathered in the SD. The job is supposed to produce
new data files  in the  SD as well. GM does not guarantee the
availability of any other places accessible by the job other than
SD (unless such a place is part of a requested Runtime Environement).
The SD is also the only place which is controlled by the GM. It is
is accessible by the user from outside through HTTPS and GridFTP protocols.
Any file created outside the SD is not controlled by the GM. Any exchange
of data between client and GM (including also program modules) is
done through GridFTP \cite{gridftp} or HTTPS protocols depending
on the selected interface. A URL for accessing input/output files
is obtained through the Information System or the WS Information Interface
of A-Rex.

@@

  Jobs are initiated and controlled through GFS or WS. Complete
job descriptions   (JD)   are passed to the GM either
through GFS in RSL \cite{rsl} or through WS in JSDL \cite{jsdl}
coded description     . Input data files
and job executables are transferred separately through the same interface,
as described in the next section.

@@

The main task of the GM is to take care of processing input and output
data (files) of the job. Input files are gathered in the SD. There
are two ways to put a file into the SD:

\begin{itemize}
\item Download is initiated by the GM --  This is the case for
files     defined in the JD (with name and source).
      The GM alone is responsible to make
sure that all required files will be available in the SD.\\
The supported protocols for sources  at the moment are: GridFTP,
FTP, HTTP, HTTPS (HTTP over SSLv3) and HTTPg (HTTP over GSI). Also
some nonstandard sources are supported. Those are described below.
\item Upload is initiated by the user directly or through the User Interface
(UI). Because the SD becomes available immediately at the time of
submission of JD, UI can (and should) use that to upload data files
which are not otherwise accessible by the GM.  Examples of such
files  are the main executable of the job,   the
job's input files, etc. These files can (and should)
also be specified in the JD.
\end{itemize}
\textbf{There is no} other reliable way for a job to obtain input
data on the CE belonging to NorduGrid. Access to AFS, NFS, FTP, HTTP
and any other remote data transport during execution of a job is
not guaranteed (at least not by GM).

Jobs should store output files in thir SD. Like input files,
output files  belong into two groups:

\begin{itemize}
\item Files which are supposed to be moved to a \emph{Storage Element} (SE)
and optionally registered in some \emph{Indexing Service} like the
Globus \emph{Replica Location Service } (RLS)  --
The GM takes care of these files. They have to be specified in the
JD. If the job fails during any stage of processing, no attept is
made to transfer those files to their final destination, unless
the  option \emph{preserve=yes} is specified in their URLs.
\item Files which are supposed to be fetched by the user -- The user has
to use a tool like the UI to obtain these files. They \textbf{must}
also be specified in the JD.
\end{itemize}

@@

\caption{\label{job states diagram}Job states}
\end{figure}
A user can examine the state of a job by querying the NorduGrid Information
System (IS) using the UI or any other suitable tool. Please remember
that the IS can manipulate  state names to make them more user
friendly and to combine them with states introduced by other parts
of the whole setup. Another way is to access virtual informational
files through GridFTP interface or to use query method of WS interface.

Configuration can put limits on the amout of simultaneous jobs in
some states. If such a limit is reached, a job ready to enter into
the state in question will stay in it's current state waiting for
a free slot. This situation is presented by prepending the prefix
\textbf{PENDING:} to the current state name  \textbf{}
in the job's status mark.

Below is the description of all actions taken by the GM at every state:

\begin{itemize}
\item \textbf{Accepted} - In this state the job has been submitted to
a CE but is not processed yet. The GM will analyze the JD and move
to the next stage. If the JD can not be processed the job will be
canceled and moved to the state \textbf{Finishing}.
\item \textbf{Preparing} - The input data is being gathered in the SD (stage-in).
The GM is downloading the files specified in the JD and is waiting
for files which are supposed to be uploaded by the UI. If
all files are successfully gathered the job moves to the next state.
If \textbf{any} file can't be downloaded or it takes the UI too long
to upload a file,  the job moves to \textbf{Finishing} state. It
is possible to put a limit on the number of simultaneous \textbf{Preparing}
jobs. If this limit is exceeded, jobs ready to enter the \textbf{Preparing}
state      will stay in the \textbf{Accepted}
state, but prefixed with the PENDING: mark. Exceptions are jobs which
have no files to be downloaded. These are processed out of limits.
\item \textbf{Submitting} -      The job is being
passed for execution to the \emph{Local Resource Management System}
(LRMS). At the moment PBS is supported best but correspoding backends
for many other LRMSs are provided with the default installation. 
      If the local job submission
is successful the job moves to the \textbf{InLRMS} state. Otherwise
it moves to \textbf{Finishing}. It is possible to limit the agregate
number of jobs in \textbf{Submitting} and  \textbf{InLRMS}
states.
\item \textbf{InLRMS} - The job is queued or being executed in the LRMS.
The GM takes no actions except waiting until the job finishes.
\item \textbf{Cancelling} - Necessary action to cancel the job in the LRMS
is being taken.
\item \textbf{Finishing} - The output data is being processed (stage-out).
Specified data files are moved to the specified SEs and are optionally
registered at  an Indexing Service. The user can download data files
from the SD by using the UI or  other adequate tool. All the files
not specified as output files are removed from the SD at very beginning
of this state. It is possible to limit number of simultaneous jobs
in this state.
\item \textbf{Finished} - No more processing is performed by the GM. The
user can continue to download data files from the SD. The SD is kept
available for some time (default is 1 week). After that the job
is moved to the state \textbf{Deleted}. The 'deletion' time can be
obtained by querying the NorduGrid Information System  attribute
\texttt{nordugrid-pbs-job-sessiondirerasetime} of a \texttt{nordugrid-pbs-job}
entry. If a job was moved to \textbf{Finished} because of failure,
it may be restarted on request of a client. When restarted, a job
is moved to the state previous to the one in which it failed and
is assigned mark PENDING. This is needed in order to not break the
configuration limits. Exception is a job failed in \textbf{InLRMS}
state and lacking input files specified in JD. Such a job is treated
like failed in \textbf{Preparing} state.
\item \textbf{Deleted} - The job is moved to this state if the user have
not requested job to be cleaned before the SD's lifetime expires.
Only minimal subset of information about such job is kept. The SD
is not available anymore.
\end{itemize}
In  case of a failure special processing is applied to output
files. All specified output files are treated as \textbf{downloadable
by the user}. No files will be moved to their destination SE.

@@

The GM and it's components support the following data transfer protocols
and corresponding URLs: \emph{ftp, gsiftp, http, httpg, https, se,
rc} and \emph{rls.} For more information please see {}``Protocols,
Uniform Resource Locators (URL) and extensions supported in ARC''
document \cite{urls}.

@@

For each local UNIX user listed in the GM configuration a \textit{control
directory} exists. In this directory the GM stores information about
jobs belonging to that user. Multiple users can share the same \textit{control
directory}. To make it easier to recover in  case of failure, the
GM stores most information in files rather than in memory. All files
belonging to same the job have names starting with \textbf{job.ID.,}
where ID is the job identifier.

@@

\begin{itemize}
\item \textit{job.ID.status} - current state of the job. This is a plain
text file containing a single word   representing the current
state of the job. Possible values are :

\begin{itemize}
\item ACCEPTED
\item PREPARING
\item SUBMITTING
\item INLRMS
\item FINISHING
\item FINISHED
\item CANCELING
\item DELETED
\end{itemize}
\end{itemize}
See section \ref{Section: Job Flow} for a description of the various
states. Additionally each value can be prepended  the prefix
{}``PENDING:'' (like PENDING:ACCEPTED, see section \ref{Section: Job Flow}).
This is used to show that a job  is \emph{ready} to be moved
to the next state but it has to stay in it's current state \emph{only}
because otherwise some limits set in the configuration would be
exceeded.

\begin{itemize}
\item \textit{job.ID.description} - contains the RSL description of the
job.
\item \textit{job.ID.local} - information about the job used by the GM.
It consists of lines of format \textit{{}``name = value''} . Not
all of them are always available. The following names are defined:

\begin{itemize}
\item \textit{subject} - user certificate's subject, also known as the distinguished
name (DN)
\item \textit{starttime} -  GMT time when the job was accepted represented
in the Generalized Time format of LDAP 
\item \textit{lifetime} - time period to  preserve the SD after the
job has finished in seconds
\item \textit{cleanuptime} -  GMT time when the job should be removed
from the cluster and it's SD deleted in Generalized Time format
\item \textit{notify} - email addresses and flags to send mail to about
the job specified status changes
\item \textit{processtime} -  GMT time when to start processing the job
in Generalized Time format
\item \textit{exectime} -  GMT time when to start job execution in Generalized
Time format
\item \textit{expiretime} -  GMT time when the credentials delegated
to the job expire in Generalized Time format
\item \textit{rerun} - number of retries left to rerun the job
\item \textit{jobname} - name of the job as supplied by the user
\item \textit{lrms} - name of the LRMS backend     to be used
for local submission
\item \textit{queue} - name of the queue to run the job at
\item \textit{localid} - job id in LRMS (appears only after the job
 has reached state \textbf{InLRMS})
\item \textit{args} - executable name followed by a list of command-line
arguments   
\item \textit{downloads} - number of files to download into the SD before
execution
\item \textit{uploads} - number of files to upload from the SD after execution
\item \textit{gmlog} - directory name which holds files containing information
about the job when accessed through GridFTP interface
\item \textit{clientname} - name (as provided by the user interface) and
ip address:port of the submitting client machine

\item \textit{clientsoftware} - version of software used to submit the job
\item \textit{sessiondir} - the job's SD  
\item \textit{failedstate} - state in which job failed (available only
if it is possible to restart the job)
\item \textit{jobreport} - URL of a user requested \emph{logger service}.
The GM will also send job records to this service in addition to the
default logger service configured in the configuration

\end{itemize}
\end{itemize}
This file is filled partially during job submission and fully when
the job moves from the \textbf{Accepted} to the \textbf{Preparing}
state.

@@

The job submission and control interface is based on document produced
by the OGF OGSA Basic Execution Services (BES) Working Group\cite{bes}.
The exchange of SOAP messages is performed over the HTTPS protocol.
The BES interface is represented by two port-types - BES-Management
and BES-Factory. The former is made to control the A-REX service itself
and thus defines operations to start and stop the functionality of
the BES service. A-REX does not implement remote control of service
functionality. Hence the BES-Management port-type is not implemented.
The BES-Factory port-type provides operations to submit new jobs (to
create an activity in terms of BES) and to monitor its state. It also
has an ability to provide information about the service. A-REX fully
implements the functionality of this port-type.

@@

\begin{itemize}
\item \emph{NotAuthorizedFault}: Indicates that the client is not allowed
to do this operation.
\item \emph{InvalidActivityIdentifierFault}: There is no such job/activity.
\item \emph{CantApplyOperationToCurrentStateFault}: The requested transition
is not possible.
\end{itemize}
\end{itemize}
\end{itemize}
On result of this command, the job should be put into the requested
state. If such a procedure cannot be performed immediately then the
corresponding sequence is initiated and fault OperationWillBeAppliedEventuallyFault
will be returned.

@@

\begin{itemize}
\item \emph{DelegatedTokenType DelegatedToken}: Contains an X509 proxy certificate
based on the public key from the DelegateCredentialsInit signed by
the user's proxy certificate. Also includes the Id element which identifies
the private key stored at the service side associated with these credentials.
The reference element refers to the object to which these credentials
should be applied  in a way specific to the service. The same element
must also be used for delegating credentials as part of other operations
on service.
\end{itemize}
\item Output(s)

@@

\begin{lyxlist}{00.00.0000}
\item [{Filename}] - name of the data file on the executing node.
\item [{\emph{Source}}] - source where the file will be taken from before
execution.
\item [{\emph{Target}}] \emph{-} destination the file will be delivered
to after execution.
\end{lyxlist}
\item [{\emph{Input}\ (POSIX)}] - file to be used as standard input for
the executable.
\item [{\emph{Output}\ (POSIX)}] - file to be used as standard output
for the executable.
\item [{\emph{Error}\ (POSIX)}] - file to be used as standard error for
the executable.
\item [{\emph{MemoryLimit}\ (POSIX)}] - amount of physical memory needed
for execution.
\item [{\emph{TotalPhysicalMemory}}] - same as MemoryLimit.
\item [{\emph{IndividualPhysicalMemory}}] - same as MemoryLimit.
\item [{\emph{CPUTimeLimit}\ (POSIX)}] - maximal amount of CPU time needed
for execution.
\item [{\emph{TotalCPUTime}}] - same as CPUTimeLimit.
\item [{\emph{IndividualCPUTime}}] - same as CPUTimeLimit.
\item [{\emph{WallTimeLimit}\ (POSIX)}] - amount of clock time needed
for execution.
\item [{\emph{TotalCPUCount}}] \emph{-} number of CPUs needed for execution.
\item [{\emph{IndividualCPUCount}}] - same as TotalCPUCount.
\end{lyxlist}

@@

\begin{itemize}
\item \textit{list} - stores names of the files (8 digit numbers) and corresponding
URLs delimited by blank space. Each pair is delimited by some amount
of \textbackslash{}0 codes. Also creation and expiration times are
stored if available
\item \textit{old} - stores URLs which have been removed from cache. Records
are delimited by some amount of \textbackslash{}0 codes and are meant
to be removed by some external routine.
\item \textit{new} - stores URLs which have been added to cache. Records
are delimited by some amount of \textbackslash{}0 codes and are removed
when corresponding files are removed from cache. They can also be
handled by some external routines. Every time record is added to \emph{old}
it is removed from \emph{new}.
\item \textit{statistics} - consists of strings containing \emph{name=value}
pairs. Following names are defined:

@@

If a job requests an input file which is subject  for
cacheing, it is stored in the cache directory instead of the SD.
The file is made available to the job by either  soft-linking it
  in the SD         
    or copying it to the SD. The latter option
is more secure and hence advised.

Before downloading a file the GM tries to determine it's size and
then tries to preallocate space in the cache directory by writing
a file of the same size. If this fails (possibly because the file
system has no more space), it tries to the remove oldest cached files
which are not being used by any job. This means that \textbf{the
hard limit of cache size is the space available in the file-system}.
In case cache gets full and it is impossible to free enough space,
the download fails and then is retried without using cache.

Before giving access to an already cached file, the GM contacts the
original file source to check if the user     has
suficient access rights to the file. Not all protocols  
 make this check possible.

Also, file creation or validity times are checked to make sure the
cached file is fresh . If it is impossible to obtain creation
and invalidation times  , the file is invalidated 24 hours
after it was downloaded.

 The GM checks the cache periodically. If  the space used
by cache exceeds the high water-mark given in the configuration file
(\emph{softsize}), it tries to remove the oldest unused files until
cache size drops below the  low water-mark. This sets the
soft limit of cache size.

There are two kinds of caches supported. Files in \textit{private}
cache are owned by the Unix user to which a grid user is mapped. Those
files are readable only by that particular Unix user. Another kind
of cache is \textit{shared}. Files in the shared cache are owned by
the Unix user who started GM and are readable by everyone.

@@

The GM consists of several separate executable modules. These
are:

\begin{itemize}
\item \textit{grid-manager} - The main module. It is responsible for processing
 jobs, moving them through states and running other modules.
\item \textit{downloader} - This is a module responsible for gathering input
files in the SD. It processes the \textit{job.ID.input} file and updates
it.
\item \textit{uploader} - This module is responsible for delivering output
files to the specified SEs and registration at  an Indexding
Service (like RLS) as needed. It processes and updates the \textit{job.ID.output}
file.
\item \emph{cache-register} - Utility to register cached data into an Indexing
Service    . It reads and modifies cache informational
files \emph{old} and \emph{new} (as described in section \ref{sec:cache}).
Configuration is read directly from the GM's configuration file (see
section \ref{SubSection: ConfigFile}). It is run by the GM every
5 minutes.
\item \emph{frontend-info-collector} - Utility to gather information about
the frontend  . It puts  collected information into the \emph{job.ID.diag}
file.
\item \emph{gm-kick} - Sends a signal to the GM though a FIFO file to wake
it up. It's used to increase responsiveness of GM.
\end{itemize}
The following modules are always run under the Unix account to which
a grid user is mapped.

\begin{itemize}
\item \textit{smtp-send.sh} and \textit{smtp-send} - These are the modules
responsible for sending e-mail notifications to the user. The format
of the mail messages can be easily changed by editing the simple shell
script \textit{smtp-send.sh}. 
\item \textit{submit-{*}-job} - Here {*} stands for the name of the LRMS.
Curently supported LRMS are PBS/Torque, Condor and SGE. Also \emph{fork}
pseudo-LRMS is supported for testing purposes. This module is responsible
for  job submission to the LRMS.
\item \textit{cancel-{*}-job} - This script is for canceling  jobs
which have been already submitted to the LRMS.
\item \textit{scan-{*}-job} -This shell script is responsible for notifying
the GM about completion of  jobs. It's implementation for PBS 
uses server logs to extract information about jobs. If logs are not
available it uses the less reliable \emph{qstat} command for that.
Other backends use different techniques.
\end{itemize}
 In addition, there is also an administration utility:

\begin{itemize}
\item \textit{gm-jobs} - prints a list of jobs available on the cluster
and the number of jobs in each state.\\
\hspace*{0.5cm}gm-jobs {[}-h] {[}-l] {[}-u uid] {[}-U name]\\
\hspace*{0.5cm}-l -- print more information about each job,\\
\hspace*{0.5cm}-u -- pretend utility is run by user with id \emph{uid},\\
\hspace*{0.5cm}-U -- pretend utility is run by user with name \emph{name}.
\end{itemize}
GM comes with plugins useable for various authorization purposes (see
for example the description of \emph{authplugin} command below):
